---
title: "APA table demo"
author: "Cylcia Bolibaugh"
date: "23/09/2021"
output: html_notebook
      
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE,  message=FALSE, warning=FALSE)
# install any missing packages first
library("papaja")
library("tidyverse")
library("here")
library("flextable")
```

```{r load_data}
# Load in the GTA_data.csv dataset
GTA_data <- read_csv(here("example_GTA_data.csv")) # have used 'here' package to make relative path loading work

```

```{r set_classes}
# Change variable classes
GTA_data$Q2.9_student_type <- as.factor(GTA_data$Q2.9_student_type)
GTA_data$Q2.10_language <- as.factor(GTA_data$Q2.10_language)
GTA_data$Participant <- as.factor(GTA_data$Participant)
GTA_data$deletion_type <- as.factor(GTA_data$deletion_type)
```
A friend wants to generate his APA style summary tables from r to a word output. An example of his desired output table is in the repo (example_APA analysis.xls) 

## Overall aim
In order to generate the table, I will need to extract the numerator, denominator, derive the percent, and glue these into a single cell, for each question item. I will then need to reshape the dataframe so it is in the desired format, with values displayed under nested group headings. Finally, I will use a table package to format headers etc and render in word/pdf.


### pseudocode - what needs to happen
1) extract denominators - tally the total number of cases grouped by student type, and language and assign to a df
2) extract numerators - tally number of cases for each variable when grouped by student type, and language, and assign to a df
3) derive the percentage - join the two dataframes and create a new variable (percent) using mutate
4) 'glue' all values into a single cell - concatenate the numerator, denominator and percent
5) run this workflow across all Q.10 variables
6) reshape the dataframe so it is in the desired format - pivot from wide to long to wide again  
7) send to flextable

```{r crosstabs_tally}
# Create a small df holding crosstab totals with values filtered
groupCount <- GTA_data %>% 
  filter(deletion_type == 0) %>% 
  group_by(Q2.9_student_type, Q2.10_language) %>% tally()

```

```{r occurrence_tally}
# Tally the number of occurrences for each cross-tabs cell, using purrr style notation
Q10Tally <- GTA_data %>%
  filter(deletion_type == 0) %>% 
  group_by(Q2.9_student_type, Q2.10_language) %>%
  summarise(across(starts_with("Q3.10"),  ~ sum(.x, na.rm = TRUE), .names = "{.col}"))
```

```{r output_join}
# Join dataframes, generate the final output column with proportion and percentage in brackets, with the prefix concat_, assign to new df called Q10
Q10 <- right_join(Q10Tally, groupCount) %>%
  mutate(across(starts_with("Q3.10"),
                ~paste0(.x, "/", n, " (", round(.x/n*100),"%",")"),
                .names = "concat_{.col}"))


```


```{r prop_demo}
# Demonstrate how to generate a variable called prop_ that contains the tallied response over the numerator - this step has been incorporated in previous chunk, but there's a lot going on...
Q10 <- Q10 %>%
  mutate(across(starts_with("Q3.10"),
                ~paste0(.x, "/", n),
                .names = "prop_{.col}"))


```



```{r output_reshape}
# First get data in long format, and then regroup in wide by the two grouping variables
wideQ10 <- Q10 %>% select(Q2.9_student_type, Q2.10_language, starts_with("concat")) %>%
  pivot_longer(
    cols = starts_with("concat")
  ) %>% pivot_wider(
    names_from = c(Q2.9_student_type, Q2.10_language)
    ) 


# Remove concat_ prefix with stringr package
wideQ10$name <- str_remove(wideQ10$name, "concat_")

# Remove unused item.
wideQ10 <- wideQ10 %>% filter(name != 'Q3.10_9_technology_use_count')
```


```{r create_flextable}

# Use flextable package: https://ardata-fr.github.io/flextable-book/index.html

# Create table
myft <- flextable(wideQ10)

# Change column headers
myft <- set_header_labels(myft, 
                          name = "", 
    `EU student_GTA whose English is the first language (L1 GTA)` = "L1", 
    `EU student_GTA whose English is the second language (L2 GTA)` = "L2",
    `Home student_GTA whose English is the first language (L1 GTA)` = "L1",
    `non-EU international student_GTA whose English is the first language (L1 GTA)`= "L1",
    `non-EU international student_GTA whose English is the second language (L2 GTA)` = "L2" )

# Add additional header row
myft <- add_header_row(
  x = myft, values = c("", "EU students", "Home students", "non-EU international students"),
  colwidths = c(1, 2, 1, 2))

# Set caption
myft <- set_caption(myft, "Table XXX.Proportion of respondents (%) selecting their use of each online technology type")


```


```{r format_flextable}
# set a random theme, but borders etc can be customised. Look at vignette to prettify, or just export to word 
myft <-theme_booktabs(myft)
```

```{r print_flextable}
# export by copying
print(myft, preview = "docx") # send a copy to clipboard and format there

# Or export to new doc. I've created a new .docx in the current directory.
save_as_docx("Q10_table" = myft, path = here("Q10_table.docx"))

# Or if creating a large number of tables, create them in .rmd and then knit to word instead of html. This will create a single doc with all tables in one: https://ardata-fr.github.io/flextable-book/rendering.html#r-markdown-documents
```

